---
title: "01-Data Import"
author: "lucaraess"
date: "2020-05-12"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data import
```{r}
#Normally not necessary if you work with a project (it is just because this data is from somewhere else)
setwd("C:/Users/lucar/Github/")

cells <- read.csv(file = "./ImcSegmentationPipeline/output/cpout/cell.csv", stringsAsFactors = FALSE)

image <- read.csv(file = "./ImcSegmentationPipeline/output/cpout/Image.csv", stringsAsFactors = FALSE)

relationships <- read.csv(file = "./ImcSegmentationPipeline/output/cpout/Object relationships.csv")

#panel <- read.csv(file = , stringsAsFactors = FALSE)

```

## Data cleaning
Overview of the acquired features
```{r}
unique( sub("_c[0-9]*$", "", colnames(cells)) )
```

Extract the desired data columns by name
```{r}
cur_counts <- cells[,grepl("Intensity_MeanIntensity_FullStackFiltered",
colnames(cells))]
```
Rescale image intensities
```{r}
cur_counts <- cur_counts * image$Scaling_FullStack[1]
```

## Cell-specific metadata
```{r}
library(S4Vectors)
cell_meta <- DataFrame(CellNumber = cells$ObjectNumber,
                       Center_X = cells$Location_Center_X,
                       Center_Y = cells$Location_Center_Y)
```

## Image-specific metadata
Overview acquired features
```{r}
colnames(image)[grepl("Metadata_", colnames(image))]
```

```{r}
library(stringr)

image$Metadata_acname
image$Metadata_roiid
```

Add cell number information to cell_metadata
```{r}
cell_meta$ImageNumber <- cells$ImageNumber
```

Split acquisition information
```{r}
ac_info <- str_split(image$Metadata_acname, "_", simplify = TRUE) 
cell_meta$BatchId <- ac_info[cell_meta$ImageNumber,2] 
cell_meta$SampleId <- ac_info[cell_meta$ImageNumber,3]
cell_meta$ROI <- image$Metadata_roiid[cell_meta$ImageNumber]
```

Define row names
```{r}
rownames(cell_meta) <- paste(cell_meta$SampleId, cell_meta$ROI, cell_meta$CellNumber, sep = "_")
```

## Load Feature-specific metadata
Note: All _full.csv contain the same information
```{r}
channel_mass <- read.csv("./ImcSegmentationPipeline/output/tiffs/20170906_FluidigmONfinal_SE_s0_p4_r5_a5_ac_full.csv",
header = FALSE)
```

Order pannel by isotope
```{r}
#panel information for this analysis not available
# library(DT) 
# DT::datatable(panel)
# panel <- panel[match(channel_mass[,1], panel$Metal.Tag),]
```
Change rownames to protein id
```{r}
# rownames(panel) <- panel$Target
```
Re-order column names in cur_counts to match panel info
```{r}
# channelNumber <- as.numeric(sub("^.*_c", "", colnames(cur_counts)))
# cur_counts <- cur_counts[,order(channelNumber, decreasing = FALSE)]
```

```{r}
library(igraph) 
# Construct neighbour data.frame 
# First in the ImageNumber_ObjectNumber format
cur_df <- data.frame(CellID_1 = paste0(relationships$First.Image.Number, "_", relationships$First.Object.Number),
CellID_2 = paste0(relationships$Second.Image.Number, "_", relationships$Second.Object.Number))

# Create simple cell IDs 
cellID <- paste0(cell_meta$ImageNumber, "_", cell_meta$CellNumber)

# Change cell IDs
cur_df$CellID_1 <- rownames(cell_meta)[match(cur_df$CellID_1, cellID)]

cur_df$CellID_2 <- rownames(cell_meta)[match(cur_df$CellID_2, cellID)] 

# Build graph 
g <- graph_from_data_frame(cur_df)
g
```

Create SingleCellExperiment
```{r}
library(SingleCellExperiment)

# Create SCE object 
sce <- SingleCellExperiment(assays = list(counts = t(cur_counts)))

# Set marker name as rownames and cellID as colnames
# rownames(sce) <- rownames(panel)
colnames(sce) <- rownames(cell_meta)

colData(sce) <- cell_meta 
rowData(sce) <- panel 
metadata(sce) <- list(graph = g)
sce
```

## Save SCE object

```{r}
#doesnt work at the moment (probably working directory confusion)
# saveRDS(sce, file = "./data")
```


