---
title: "01-GenerateSpillmat"
author: "lucaraess"
date: "2020-06-16"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Estimate signal spillover from single metal spots
Script according to: https://github.com/BodenmillerGroup/cyTOFcompensation/blob/master/scripts/imc_generatespillmat_long.Rmd

## load all libraries
```{r Libraries, message=FALSE}
library(CATALYST)
library(data.table)
library(ggplot2)
library(flowCore)
library(dplyr)
library(dtplyr)
library(stringr)
library(ggpmisc)

source('./code/spillover_imc_helpers.R')
```

## setup the configuration variables
```{r Setup}
# list of folders that contain each a complete single stain acquisition (e.g. in case that one wants to run and compare multiple single stains from different days)
# fols_ss = c('./data/Figure_S5/Spillover_Matrix_2','./data/Figure_S5/Spillover_Matrix_1' )
# output folder
# fol_out = './output/'
# name prefix for all output
prefix ='paper_version_'
```

## load single stains
### Data loading
```{r}
# load the data
list_img_ss <-lapply(fols_ss, load_ss_fol)
names(list_img_ss) <- fols_ss
```

### Adapt the column names to be recognized metal names by CATALYST 
CATALYST needs to have the metal names in the format (METAL)(MASS)Di
```{r}
list_img_ss = lapply(list_img_ss, function(x) lapply(x, fixnames))
dats_raw = lapply(list_img_ss, imglist2dat)
```

### Extract the single stain masses from the acquisition name
This needs to be changed in case a different naming scheme is used!
```{r Get bc masses}
for (dat in dats_raw){
  dat[, metal:= strsplit(.BY[[1]], '_')[[1]][3],by=file]
  dat[, mass:= as.numeric(str_extract_all(.BY[[1]], "[0-9]+")[[1]]),by=metal]
}
```

## Visualization of the raw data

### Calculate per-file medians
```{r}
dats_raw_sum = rbindlist(lapply(dats_raw, calc_file_medians),idcol = T)
```


### Visualize per-file medians
Plots the median of the data. It is recommended to have >200 counts for all the channels.
This is also a good plot to check if the metal spots really contain the correct metal!
```{r fig.height=13, fig.width=20}
dats_raw_sum %>%
  ggplot(aes(x=1, y=med, color=.id))+
  facet_wrap(~file+metal, scales = 'free_y')+
  geom_label(aes(label=variable), size=4)
```

###  Optional data bining

If the median per-pixel intensities are to low, it could be worth to sum up some consecuteive pixels to get a better accuracy for the estimation
(here not the case). This is valid because for segmentation based quantitative image analysis usually anyways pixels are aggregated. If the binning is choosen to big, there is however a potential accumulation of background noise.

--> Check online Code

## CATALYST based compensation

### estimate the spillover
To estimate the spillover, the (aggregated) pixel values are first debarcoded using CATALYST, treating them like single cells. This step acts as a quality filter to remove background/noisy/weak pixels as well as pixels with artefacts (e.g. specles with strong signal in many channels).
If the true metal was correctly encoded in the filename, the 'remove_incorrect_bc' option will check the debarcoding and remove events assigned to the wrong barcode.

Then this identified, strong single stain pixels will be used for the spillover estimation.

```{r Binned}
res = lapply(dats_agg, function(x) re_from_dat(x,
                                                ss_ms=x[!is.na(mass), unique(mass)],
                                                minevents = 40,
                                                correct_bc = T))
sms = lapply(res, function(x) computeSpillmat(x))
```

### save the spillover matrices

```{r}
for (i in seq_along(sms)){
  outname = file.path(fol_out, paste0(prefix, basename(fols_ss[i]),'_sm.csv'))
  write.csv(sms[[i]],file = outname)
}
```

### Visualization of the spillover matrix

```{r}
for (i in seq_along(sms)){
  print(names(dats_agg)[i])
  ss_ms = dats_agg[[i]][!is.na(mass), unique(mass)]
  p = CATALYST::plotSpillmat(ss_ms,sms[[i]])
  print(p)
}
```

--> Additional code chunks contain:
-Quality indicators
-Linearity of the spillover
